\textbf{universal corpus}

### 2014-04-25 docs/computEL-final/cluster.tex
Updated results, added discussion

We tested three methods, which differ in how they define the distance between two clusters. The \texttt{single} method considers all pairs of languages with one from each cluster, and finds the smallest distance. The \texttt{complete} method instead finds the largest distance. The \texttt{weighted} method calculates an average across all the pairwise differences, weighted according to the languages' positions in the tree.

@@

    \begin{tabular}{l|ccc}
    ~        & Precision & Recall       & F-score    \\ \hline
    \texttt{single} & 0.1958	& 0.6755	 & 0.1240  \\
	\texttt{complete} & \textbf{0.3153}	& 0.1699	 & \textbf{0.1420} \\
	\texttt{weighted} & 0.0614	& \textbf{0.8565}	 & 0.1099 \\ \hline
	\emph{random} & 0.1925 &	0.0927 & 	0.0692 \\
    \end{tabular}
\caption{Comparison of clustering algorithms}
\label{table:cluster}
\end{centering}
\end{table}

@@

In table \ref{table:cluster}, we give results of our clustering experiments, comparing the three clustering methods to a random baseline (clustering based on random distances, averaged over several runs).  The F-scores are comparable to those reported by \newcite{georgi2010wals}, even though we have only used surface features, while they used typological features taken from WALS.  This demonstrates that it possible for cross-linguistic research to be conducted even based on extremely shallow features.

@@

It is worth noting that precision is higher than recall.  This is perhaps to be expected, given that related languages using wildly differing orthographies will appear different.  Nonetheless, our system is reasonably capable of identifying languages which are both related and written similarly.

\paragraph{Qualitative Inspection.}

>>>>

We tested four methods, which differ in how they define the distance between two clusters. The \texttt{complete} method considers all pairs of languages with one from each cluster, and finds the largest distance. The \texttt{single} method instead finds the smallest distance. The \texttt{average} and \texttt{weighted} methods calculate averages across all the pairwise differences, either unweighted, or weighted according to the languages' positions in the tree.

@@

    \begin{tabular}{l|ccc}
    ~        & Precision & Recall       & F-score    \\ \hline
	\texttt{complete} & \textbf{0.3142}	& 0.1938	 & \textbf{0.1556} \\
	\texttt{single} & 0.1949	& \textbf{0.7850}	 & 0.1282 \\
	\texttt{average} & 0.2216	& 0.5163	& 0.1348 \\
	\texttt{weighted} & 0.2326	& 0.3343	& 0.1271 \\ \hline
	\emph{random} & 0.1836	& 0.0917	& 0.0675 \\
    \end{tabular}
\caption{Comparison of clustering algorithms}
\label{table:cluster}
\end{centering}
\end{table}

@@

In table \ref{table:cluster}, we give results of our clustering experiments, comparing the three clustering methods to a random baseline (clustering based on random distances, averaged over 20 runs).  It is worth noting that precision is higher than recall.  This is perhaps to be expected, given that related languages using wildly differing orthographies will appear different.  Nonetheless, our system is reasonably capable of identifying languages which are both related and written similarly.

To allow a closer comparison with \newcite{georgi2010wals}, we can also calculate pairwise scores - i.e. considering when pairs of languages are in the same cluster or the same class. For 147 clusters, the \texttt{complete} method gives a pairwise f-score of 0.147, while they report 0.140. The figures are not quite comparable since we are evaluating over a different set of languages; nonetheless, this result is suggestive, given that we have only used surface features, while they used typological features taken from WALS.  This demonstrates that it possible for cross-linguistic research to be conducted even based on extremely shallow features.

@@

\paragraph{Qualitative Inspection.}
Ga: It's actually pretty difficult to interpret the clusters directly because there's so much noise. I will try to write something better, but here are a few thoughts I've had:

Using the \texttt{complete} method with 147 clusters, we obtain a mix of large and small clusters. There are 27 singleton clusters, while the largest cluster contains 123 languages. Although these singleton clusters belong to classes of various size, it is interesting that the distribution of sizes is similar to the distribution in the gold standard data, where we have 27 single classes, and the largest class contains 150 languages.  This isn't true for the other clustering methods - the \texttt{single} method produces 138 singleton clusters, with almost all languages in a single cluster of size 968.

### 2014-04-25 docs/computEL-final/intro.tex
Updated results, added discussion

We believe the resulting corpus, which at present covers 1347 languages from 106 language families, is the first of its kind: large enough and consistent enough to allow broadly multilingual language processing. To test this claim, we use SeedLing in a sample application (section \ref{sec:cluster}): the task of language clustering. With no additional pre-processing required, we extract surface-level features (frequencies of character n-grams and words) for estimating the similarity of two languages. Unlike most previous approaches to the task, we make no use of resources curated for aims of linguistic typology (e.g. values of typological features as in WALS \cite{wals}, Swadesh word lists). Despite our approach being highly dependent on orthography, we achieve a reasonable clustering performance over the languages in our dataset, demonstrating SeedLing's utility in cross-linguistic research.

>>>>

We believe the resulting corpus, which at present covers 1347 languages from 106 language families, is the first of its kind: large enough and consistent enough to allow broadly multilingual language processing. To test this claim, we use SeedLing in a sample application (section \ref{sec:cluster}): the task of language clustering. With no additional pre-processing required, we extract surface-level features (frequencies of character n-grams and words) for estimating the similarity of two languages. Unlike most previous approaches to the task, we make no use of resources curated for aims of linguistic typology (e.g. values of typological features as in WALS \cite{wals}, Swadesh word lists). Despite our approach being highly dependent on orthography, our clustering performance matches the results obtained by \newcite{georgi2010wals} using typolological features, which demonstrates SeedLing's utility in cross-linguistic research.
